---
title: "DATA 605 WK 3 Assignment"
author: "Jun Pan"
date: "February 17, 2019"
output: html_document
---

```{r}
require("pracma")
```


Problem 1.
(1)What is the rank of the matrix A? 
```{r}
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow=4, byrow=TRUE)
A
```
```{r}
#Change a Matrix Into its Echelon Form
rref(A)
```
The maximum number of linearly independent vectors in a matrix is equal to the number of non-zero rows in its row echelon matrix. Therefore, to find the rank of a matrix, we simply transform the matrix to its row echelon form and count the number of non-zero rows.

There are 4 pivot rows, so the rank of $A$ is 4. 
```{r}
Rank(A)
```
double check by r programing which shows Rank=4.

reference:
        https://stattrek.com/matrix-algebra/echelon-transform.aspx#MatrixA
        https://stattrek.com/matrix-algebra/matrix-rank.aspx

(2)Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero? 

Anwser: maximum rank is n, minum rank is 1.

reference: https://stattrek.com/matrix-algebra/matrix-rank.aspx

(3) What is the rank of matrix B? 
```{r}
B <- matrix(c(1,2,1,3,6,3,2,4,2), nrow=3, byrow=TRUE)
B
```
```{r}
#change matrix B into echelon form
rref(B)
```
Accordint to the echelon form of matrix B, we conclude the rank of B is 1.

```{r}
#Double check with R programming
Rank(B)
```

Problem 2
Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution
```{r}
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow=3, byrow=TRUE)
A
```

Because $A$ is a "upper triangular matrix"", its eigenvalues are values on the diagonal, so $\lambda_1=1$, $\lambda_2=4$ and $\lambda_3=6$.

reference:http://mathonline.wikidot.com/determining-eigenvalues-from-upper-triangular-matrices-of-li
          https://math.stackexchange.com/questions/1971598/eigenvectors-of-a-triangular-matrix
          
          
```{r}
# Double-check eigenvalues in R programming
eigen(A)$values
```         

The "characteristic polynomial"" is 
$p_A(\lambda) = (1-\lambda)(4-\lambda)(6-\lambda)$ or
$p_A(\lambda) = -\lambda^3+11\lambda^2-34\lambda+24$.

If $\lambda=1$, then $A - 1I_3$ is row-reduced to  


```{r}
rref(A - 1 * diag(3))
```
 
$$
 \begin{bmatrix}
 0 &1 &0\\
 0 &0 &1\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$

Then $v_1=v_1$ and $v_2=0$ and $v_3=0$. The "eigenspace" is            
          
$$
 E_{\lambda=1}=
 \Bigg\langle \Bigg\{
 \begin{bmatrix}
 1\\
 0\\
 0
 \end{bmatrix}
 \Bigg\} \Bigg \rangle
$$          
If $\lambda=4$, then $A - 4I_3$ is row-reduced to 
```{r}
rref(A - 4 * diag(3))
```          
$$
 \begin{bmatrix}
 1 &-\frac{2}{3} &0\\
 0 &0 &1\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$          

Then $v_1 - \frac{2}{3}v_2=0$ and $v_3=0$. 

Or $v_1=v_1$ and $v_2=\frac{3}{2}v_1=1.5v_1$ and $v_3=0$. 

The "eigenspace" is  

$$
 E_{\lambda=4}=
 \Bigg\langle \Bigg\{
 \begin{bmatrix}
 1\\
 1.5\\
 0
 \end{bmatrix}
 \Bigg\} \Bigg \rangle
$$
Finally, if $\lambda=6$, then $A - 6I_3$ is row-reduced to 

```{r}
rref(A - 6 * diag(3))
```

$$
 \begin{bmatrix}
 1 &0 &-1.6\\
 0 &1 &-2.5\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$

Then $v_1-1.6v_3=0$ and $v_2-2.5v_3=0$. 

Or $v_1=1.6v_3$ and $v_2=2.5v_3$ and $v_3=v_3$. 

The "eigenspace" is  

$$
 E_{\lambda=6}=
 \Bigg\langle \Bigg\{
 \begin{bmatrix}
 1.6\\
 2.5\\
 1
 \end{bmatrix}
 \Bigg\} \Bigg \rangle
$$














          
          
          